---
title: "Untitled"
author: "Long Nguyen"
date: "10/30/2017"
output:
  html_document:
    keep_md: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(scales)
```

## 27.2.1  
### 1   
### 2   
### 3   
Both can run R-script
R notebook cannot knit like R markdown
### 4   
I think the only different is file output and how the file is format, pdf have margin, html doesn't have the same margin

## 27.3.1  
### 1   

#### Long Nguyen
Long Nguyen 
caycham1@gmail.com  

##### Education  
 * California Institute of Technology
 * B.S. Mechanical Engineering

##### Work Experience  
  * Red Ventures Software Engineer
  * 2016-Present
  * Build CLI tool for Atom platform
  * Add logging functionality to track errors

### 2  
foooooootttttnoooooooottteeeee [^1]
[^1]: Here is the footnote.  
***  
> block quote

### 3  

## 27.4.7  
### 1  
```{r}
knitr::kable(diamonds %>% 
  group_by(cut) %>% 
  summarise(avg_carat = mean(carat)))
knitr::kable(diamonds %>% 
  group_by(color) %>% 
  summarise(avg_carat = mean(carat)))
knitr::kable(diamonds %>% 
  group_by(clarity) %>% 
  summarise(avg_carat = mean(carat)))

```
### 2   
```{r}
knitr::kable(diamonds %>% 
  arrange(desc(carat)) %>% 
  top_n(20))
```

### 3  
```{r}
total <- diamonds %>%
  count()
bigger <- diamonds %>% 
  filter(carat >= 2.5) %>% 
  count()
pct <- bigger / total
paste(pct)
```
