---
title: "Untitled"
author: "Long Nguyen"
date: "10/22/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(nycflights13)
make_datetime_100 <- function(year, month, day, time) {
  make_datetime(year, month, day, time %/% 100, time %% 100)
}
flights_dt <- flights %>% 
  filter(!is.na(dep_time), !is.na(arr_time)) %>% 
  mutate(
    dep_time = make_datetime_100(year, month, day, dep_time),
    arr_time = make_datetime_100(year, month, day, arr_time),
    sched_dep_time = make_datetime_100(year, month, day, sched_dep_time),
    sched_arr_time = make_datetime_100(year, month, day, sched_arr_time)
  ) %>% 
  select(origin, dest, ends_with("delay"), ends_with("time"))
```
##16.2.4  
###1  
```{r invalid string}
ymd(c("2010-10-10", "bananas"))
```
we get an error fail to parse

###2  
tzone tell us what time zone the date is in, otherwise we will have today() giving different date even though it is called at the same moment.

###3  
```{r parse date}
d1 <- "January 1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)")
d5 <- "12/30/14" # Dec 30, 2014

mdy(d1)
ymd(d2)
dmy(d3)
mdy(d4)
mdy(d5)
```

##16.3.4  
###1  
```{r departure time between dates}
flights_dt %>%
  mutate(time = hour(dep_time) * 100 + minute(dep_time),
         week = as.factor(week(dep_time))) %>%
  ggplot(aes(x = time, group = week, color = week)) +
  geom_freqpoly(binwidth = 100)
```
Using the departure time within the day graph, I don't see any deviation between dates, except for week 53, I guess no one want to fly much between christmas and new year

###2  
```{r compare dep_time}
flights_dt %>%
  mutate(dep_time2 = sched_dep_time + dep_delay * 60) %>%
  filter(dep_time2 != dep_time) %>%
  select(dep_time2, dep_time, sched_dep_time, dep_delay)
```
dep_time2 is sched_dep_time + dep_delay. If they are consistent, this would be the same number. But the date is off by 1. I think these are flight that delay over night. dep_time doesn't account for this, hence the inconsistency

###3  
```{r airtime}
flights_dt %>%
  mutate(flight_duration = as.numeric(arr_time - dep_time),
         air_time) %>%
  select(origin, dest, flight_duration, air_time)
```
I think this is caused by flying to a different timezone, or fly overnight

###4  
```{r delay}
flights_dt %>%
  mutate(sched_dep = hour(sched_dep_time)) %>%
  group_by(sched_dep) %>%
  summarise(dep_delay = mean(dep_delay)) %>%
  ggplot(aes(y = dep_delay, x = sched_dep)) +
    geom_line()
```
I think earlier delay will compound the effect and cause more flight delay later in the day

###5  
```{r week delay}
flights_dt %>%
  mutate(weekday = wday(sched_dep_time)) %>%
  group_by(weekday) %>%
  summarise(arr_delay = mean(arr_delay, na.rm = TRUE))
```
Saturday is clearly the winner here, arrive earlier on average

###6  
```{r diamond and flights}
ggplot(diamonds, aes(x = carat)) + 
  geom_density()
ggplot(flights_dt, aes(minute(sched_dep_time))) + 
  geom_density()
```
I guess they peak at whole number, as most flight depart at 9:30, 10:00 same as diamond usually have whole number carat

###7  
I am not sure I understand this question, early departure by definition is cause by flight leaving early...

##16.4.5  
###1  
Because each month may have different date, dmonth() won't make much sense since it can give different number depends on which month is included

###2  
overnight indicate if the flight is overnight. It return true if it is, and false if it isn't. This is useful to calculate arrival time of a flight

###3  
```{r month}
ymd("2015-01-01") + months(0:11)
ymd("2017-01-01") + months(0:11)
```

###4  
```{r birthday}
age <- function(birthday) {
  (birthday %--% today()) %/% years(1)
}
age(mdy("2-11-1987"))
```

###5  
```{r day}
(today() %--% (today() + years(1)) / months(1))
```
This will always return 12, because we add a year and divide by 1 month. I am not sure what this suppose to do in the first place to fix it.