---
title: "r4ds_wrangle_exercises_14-15"
author: "Long Nguyen"
date: "10/16/2017"
output:
  html_document: default
  pdf_document: default
  word_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("htmlwidgets")
library(stringr)
```
##14.2.5  
###1  
The function paste() separate string with space, paste0() does not. paste0() behave like str_c() but str_c() does not put NA together with string like paste0 does; it will just display NA.
###2  
sep is what get insert in between arguments of str_c when it is put together. collapse is what used to combines input vector into a single string (like comma)  
###3  
```{r string}
string <- "abcdef"
len <- str_length(string)
mid <- ceiling(len/2)
str_sub(string, mid, mid)
```
###4  
str_wrap() add linebreak after certain number of character
###5  
str_trim() get rid of whitespace in a string
###6  
```{r funtion}
func <- function(str, comma = ", ", and = ", and ") {
  if( length(str) > 1) {
    str_c(str_c(str[-length(str)], collapse = comma), str[length(str)], sep = and)
  } else {
    str
  }
}
func(c("a", "b", "c"))
```
##14.3.1.1  
###1  
\ will escape next character, which is nothing so nothing get display
\\ the first \ escape to show the second \ but the second \ is to escape for regular expression.
\\\ similar to the above, the second \ will escape for regular expression, the third one only escape the next character, we need 1 more if we want to display \ 
###2  
```{r match}
x<-"\"'\\"
writeLines(x)
str_view(x,'"\\\'\\\\')
```
###3  
it would match ".a.b.c"  
##14.3.2.1  
###1  
```{r literal string}
str_view("$^$", "^\\$\\^\\$$")
```
###2  
- ^y
- x$
- .{3}
- .{7,}
##14.3.3.1  
###1  
- "^[aeiou]"
- I am not sure about this one
- "^ed$|[^e]ed$"
- "i(ng|se)$"
###2  
```{r i before e}
sum(str_detect(stringr::words, "(cei|[^c]ie)"))
sum(str_detect(stringr::words, "(cie|[^c]ei)"))
```
###3  
```{r q u}
sum(str_detect(stringr::words, "q[^u]"))
```
###4  
I only know colour/color and finalize/finalise
- "ou|ise^"

###5  
- "\\(\\d{3}\\)\\d{3}-\\d{4}"
##14.4.2  
###1  
```{r str_detect}
#start or end with x
stringr::words[str_detect(words, "^x|x$")]
stringr::words[str_detect(words, "^x")|str_detect(words, "x$")]
#Find all words that start with a vowel and end with a consonant.
stringr::words[str_detect(words, "^[aeiou].*[^aeiou]$")]
stringr::words[str_detect(words, "^[aeiou]") & str_detect(words, "[^aeiou]$")]
#Are there any words that contain at least one of each different vowel?
stringr::words[str_detect(words, "a") & str_detect(words, "e") & str_detect(words, "i") & str_detect(words, "o") & str_detect(words, "u")]
```
###2
I am not sure about most vowel, but proportionally, 'a' has 100% and should be highest
##14.4.6.1  
###1  
```{r split}
str_split(c("apples, pears, and bananas"), ", +(and +)?")[[1]]
```
###2  
boundary("word") also split on punctuation
###3  
```{r split empty string}
str_split("abc def", "")[[1]]
```
##15.4.1  
###1  
Depends on if we care about these outlier value or not, the mean might be a decent metric. 
###2  
Marital and race are principal, the rest seems arbitrary
###3  
Because Not applicable got assign a value of 0
