---
title: "r4ds_wrangle_exercises_10-11"
author: "Long Nguyen"
date: "10/2/2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## 10.5

###1
Tibbles print out less data than what is shown, tibble will also show type like dbl
###2
data.frame will function like autocomplete, where `df$x` will result in `df$xyz` which can select the wrong column in some case
tibble is more explicit
###3
We need to use df[[var]] here because df$var will look for column called var
###4
```{r annoying}
annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)
annoying$`1`
ggplot(annoying, aes(x = `1`, y = `2`)) +
  geom_point()
annoying$`3` <- annoying$`2` / annoying$`1`
annoying <- rename(annoying, one = `1`, two = `2`, three = `3`)
annoying$`three`
annoying$`two`
annoying$`one`
```
###5
This function convert a vector to a data frame
###6
n_extra inside print.tbl_df control the extra columns
##11.2.2
###1
read_delim()
###2
They have: file, col_names, col_types, locale, na, comment, trim_ws, skip, n_max, guess_max, progress
###3
col_positions, without this we have no idea where the column end
###4
```{r read_delim}
read_delim("x,y\n1,'a,b'" , ",", quote = "'")
```
###5
```{r identify error}
read_csv("a,b\n1,2,3\n4,5,6")
#There are only 2 column a and b but each row that follow has 3, last column is missing
read_csv("a,b,c\n1,2\n1,2,3,4")
#There are 3 column but again the column in the data doesn't match the number of column
read_csv("a,b\n\"1")
#There is an extra " 
read_csv("a,b\n1,2\na,b")
#There is a mix of number and string in the data, the column type become string
read_csv("a;b\n1;3")
#Use , instead of ; for read_csv
```

##11.3.5
###1
I think encoding is most important because it dictate how locale is read
###2
It will throw error if decimal mark and group mark are set to the same
If decimal mark is set to , the group mark is set to .
If decimal mark is set to . the group mark is set to ,
###3
Different place can have different date/time format, these options tell use how to parse them
###4
###5
read_csv2 use ; instead of ,
###6
UTF-8 seems to be the current standard
For Asia, Japanese and Chinese has their own script. Vietnamese have ISO standard
###7
```{r parse date}
d1 <- "January 1, 2010"
parse_date(d1, "%B %d, %Y")
d2 <- "2015-Mar-07"
parse_date(d2, "%Y-%b-%D")
d3 <- "06-Jun-2017"
parse_date(d3, "%D-%b-%Y")
d4 <- c("August 19 (2015)", "July 1 (2015)")
parse_date(d4, "%B %d (%Y)")
d5 <- "12/30/14" # Dec 30, 2014
parse_date(d5, "%m/%d/%y")
t1 <- "1705"
parse_time(t1, "%H%M")
t2 <- "11:15:10.12 PM"
parse_time(t2)
```




