---
title: "r4ds_wrangle_exercises_12-13"
author: "Long Nguyen"
date: "10/9/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("babynames")
install.packages("nasaweather")
install.packages("fueleconomy")
install.packages("maps")
library("maps")
library("fueleconomy")
library("nasaweather")
library("babynames")
library(tidyverse)
library("nycflights13")
```
##12.2.2.1  
###1  
Table 1 has year, country, with variable cases and population
Table 2 has year, country, cases and population are combined to one column, there is a numeric count variable
Table 3 has year, country, cases and population represented by rate (cases/population)
Table 4a and 4b are split into year, country; 4a has cases and 4b has population
  
###2   
```{r tidy}
#table2
cases <- filter(table2, type == "cases")[["count"]]
country <- filter(table2, type == "cases")[["country"]]
year <- filter(table2, type == "cases")[["year"]]
population <- filter(table2, type == "population")[["count"]]
table2_tidy <- tibble(country,
       year,
       rate = cases * 10000 / population )
table2_tidy
```

```{r combine}
tibble(country = table4a[["country"]],
       `1999` = table4a[["1999"]] / table4b[["1999"]] * 10000,
       `2000` = table4a[["2000"]] / table4b[["2000"]] * 10000)
```

A combination of 4a and 4b requires less operation to get to the result and easier to read

###3  
We need to isolate the row that represent each of the cases:

```{r cases}
table2 %>% 
  filter(type == "cases") %>%
  ggplot(aes(year, count)) + 
  geom_line(aes(group = country), colour = "grey50") + 
  geom_point(aes(colour = country))
```
##12.3.3  
###1  
After executing gather(), numeric value is convert to character so the column type will change after spread-gather operation.
###2  
We need quote around 1999 and 2000, because they are column name
###3  
There are 2 Phillip Woods with different ages. We need to add an identifier column
```{r additional column}
people <- tribble(
  ~name,             ~key,    ~value, ~identifier,
  #-----------------|--------|------|------
  "Phillip Woods",   "age",       45, 1,
  "Phillip Woods",   "height",   186, 1,
  "Phillip Woods",   "age",       50, 2,
  "Jessica Cordero", "age",       37, 1,
  "Jessica Cordero", "height",   156, 1
)
spread(people, key, value)
```
###4
We should use gather here:

```{r gather}
preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)
gather(preg, sex, count, male, female) %>%
  mutate(pregnant = pregnant == "yes",
         female = sex == "female") %>%
  select(pregnant, count, female)
```
  
##12.4.3  
###1  
Extra dictate what to do when there are too much data in 1 location, fill dictate what to do when there are missing data
###2  
Setting it to false will create new variable and keep the old one
###3  
Extract use regex to split columns, separate have multiple different way to split a character string. Unite is much easier because putting thing together is less ambiguous than splitting them

##12.5.1  
###1  
In spread, the fill argument specify what a value to replace NA, in complete, fill also specify this but it allow more variation across different variable
###2  
Up replace NA with next value, down replace NA with previous value

##12.6.1  
###1  
Because there are 0 in the column, I believe that the 0 cases is accounted for and na.rm will not affect the study.

###2
```{r error}
who1 <- who %>%
  gather(new_sp_m014:newrel_f65, key = "key", value = "cases", na.rm = TRUE)
glimpse(who1)
who2 <- who1 %>%
 mutate(key = stringr::str_replace(key, "newrel", "new_rel"))
who3 <- who2 %>%
  separate(key, c("new", "type", "sexage"), sep = "_")
who3
who3 %>%
  count(new)
who4 <- who3 %>%
  select(-new, -iso2, -iso3)
who5 <- who4 %>%
  separate(sexage, c("sex", "age"), sep = 1)
who5
```

We will get too few variable error  
###3  
```{r redundant}
who3 <- who %>%
  gather(new_sp_m014:newrel_f65, key = "key", value = "cases", na.rm = TRUE) %>%
  mutate(key = stringr::str_replace(key, "newrel", "new_rel")) %>%
  separate(key, c("new", "type", "sexage"), sep = "_")
select(who3, country, iso2, iso3) %>%
  distinct() %>%
  group_by(country) %>%
  filter(n()>1)
```
No data is repeated, hence we don't need both iso2 and iso3 with country.  
##13.2.1  
###1  
Path: combining origin and dest
Airport_location: combining longitude and latitude. Then we merge location into path  
###2
Origin in weather should track with faa in airpoint (same location)  
###3
year, month, day, hour, origin in weather should match with year, month, day, hour, faa in airport  
###4
Since this require special representation, we should add a column for special date and it should match the month, day and year in flights  
##13.3.1  
###1
```{r surrogate key}
flights %>% 
  arrange(year, month, day, sched_dep_time, carrier, flight) %>%
  mutate(row_number = row_number()) %>%
  glimpse()
```
  
###2  
```{r lahman}
Lahman::Batting
#Primary Key: playerId, yearID
babynames::babynames
#Primary Key: year, sex, name
nasaweather::atmos
#Primary Key: lat, long, year, month
fueleconomy::vehicles
#Primary Key: id
ggplot2::diamonds
#Primary Key: There is no unique key combination for this one
```
  
###3
See mutating_joins.xml on github  

##13.4.6  
###1  
```{r delay}
flights %>%
  group_by(dest) %>%
  summarise(delay = mean(arr_delay, na.rm = TRUE)) %>%
  inner_join(airports, by = c(dest = "faa")) %>%
  ggplot(aes(lon, lat, colour = delay)) +
    borders("state") +
    geom_point() +
    coord_quickmap()
```
  
###2  
```{r add location}
flights %>%
  left_join(airports, by = c(dest = "faa")) %>%
  left_join(airports, by = c(origin = "faa"))
```

###3  
There doesn't seems to be a relation, if anything it is trending down with age, which make no sense at all

```{r old planes}
plane_age <-
  planes %>%
  mutate(age = 2013 - year) %>%
  select(tailnum, age)
flights %>%
  inner_join(plane_age, by = "tailnum") %>%
  group_by(age) %>%
  filter(!is.na(dep_delay)) %>%
  summarise(delay = mean(dep_delay)) %>%
  ggplot(aes(x = age, y = delay)) +
  geom_line()
```

###4
I am guessing precipitation cause more delay than usual

```{r precipitation}
flights %>%
  inner_join(weather, by = c("origin" = "origin",
                            "year" = "year",
                            "month" = "month",
                            "day" = "day",
                            "hour" = "hour")) %>%
  group_by(precip) %>%
  summarise(delay = mean(dep_delay, na.rm = TRUE)) %>%
  ggplot(aes(x = precip, y = delay)) +
    geom_line()
```
  
###5  
```{r storm}
flights %>%
  filter(year == 2013, month == 6, day == 13) %>%
  group_by(dest) %>%
  summarise(delay = mean(arr_delay, na.rm = TRUE)) %>%
  inner_join(airports, by = c("dest" = "faa")) %>%
  ggplot(aes(y = lat, x = lon, size = delay, colour = delay)) +
  borders("state") +
  geom_point() +
  coord_quickmap()
```

Seems like there is a storm a bit west of Carolinas
  
##13.5.1
###1  
Some airline does not report tailnum
###2  
```{r filter}
flights%>%
  group_by(tailnum)%>%
  count() %>%
  filter(n > 100)
```
###3  
```{r combine}
fueleconomy::vehicles %>%
  semi_join(fueleconomy::common, by = c("make", "model"))
```
###4
The worse delays should correspond with stormy weather, I am not sure how to gather all the storm period to compare this to
###5
anti_join(flights, airports, by = c("dest" = "faa")) tells me about flight that have destination not listed  
anti_join(airports, flights, by = c("faa" = "dest")) tells me about the airports that is not connected to NYC in 2013 
###6  
There might be multiple airline using the same plane
```{r distinct planes}
flights %>%
  group_by(tailnum, carrier) %>%
  count()
```

