---
title: "r4ds_wrangle_exercises_14-15"
author: "Long Nguyen"
date: "10/16/2017"
output: html_document
---

##14.2.5  
###1  
The function paste() separate string with space, paste0() does not. paste0() behave like str_c() but str_c() does not put NA together with string like paste0 does; it will just display NA.
###2  
sep is what get insert in between arguments of str_c when it is put together. collapse is what used to combines input vector into a single string (like comma)  
###3  
```{r string}
library(stringr)
string <- "abcdef"
len <- str_length(string)
mid <- ceiling(len/2)
str_sub(string, mid, mid)
```
###4  
str_wrap() add linebreak after certain number of character
###5  
str_trim() get rid of whitespace in a string
###6  
```{r funtion}
func <- function(str, comma = ", ", and = ", and ") {
  if( length(str) > 1) {
    str_c(str_c(str[-length(str)], collapse = comma), str[length(str)], sep = and)
  } else {
    str
  }
}
func(c("a", "b", "c"))
```
##14.3.1.1  
###1  
\ will escape next character, which is nothing so nothing get display
\\ the first \ escape to show the second \ but the second \ is to escape for regular expression.
\\\ similar to the above, the second \ will escape for regular expression, the third one only escape the next character, we need 1 more if we want to display \ 
###2  
```{r match}
x<-"\"'\\"
writeLines(x)
str_view(x,'"\\\'\\\\')
```
###3  
it would match ".a.b.c"  
##14.3.2.1  
```{r literal string}
str_view(c("$^$", "ab$^$sfas"), "^\\$\\^\\$$")
```
##14.3.3.1  
##14.4.2  
##14.4.6.1  
##15.4.1  
##15.5.1
